\LARGE HIV virulence evolution in structured epidemic models

### 2016-05-29 poster/hivlhs_eeid2016.Rnw
more makefile/ms tweaks

<<setup3, echo=FALSE,dev="tikz",cache=TRUE>>=
source("../R/hivFuns.R")
source("../R/LHSparms3.R")
Lspvl <- "$\\log_{10}~\\textrm{SPVL}$"
pars <- expand(ltab_mean)
dd <-with(pars,
     melt(data.frame(alpha=alpha,"transmission probability"=Beta2,
                     "duration (years)"=Duration2,check.names=FALSE),
          id.vars="alpha"))
gg_tradeoff <- ggplot(dd,aes(alpha,value))+facet_wrap(~variable,scale="free")+
    geom_line()+
    labs(x= Lspvl)
@

@@

\begin{block}{Summary}
\noindent
\hspace{-1.6cm}
Pathogens can evolve rapidly in response to changing
conditions (e.g., epidemic stage or public health interventions). 
Models of \textbf{eco-evolutionary dynamics} often
neglect important epidemiological processes, such as the formation and
dissolution of sexual partnerships. We
compared \textbf{models with a range of complexity} in their
representation of partnership dynamics and extra-partnership contact.
Simple random-mixing actually gave the closest match to the most realistic models; extra-pair contact washed out the effects
of epidemiological structure. Nevertheless, the large differences
in evolutionary dynamics among different epidemiological models
suggests that researchers trying to predict the evolution of
pathogens should proceed with caution.

@@

<<tradeoff_fig,echo=FALSE,fig.width=10>>=
print(gg_tradeoff)
@ 
  \begin{itemize}
    \item \textbf{tradeoff theory}: virulence evolution
      mediated by transmission-vs-clearance tradeoff
    \item still debated \cite{EbertBull2003,alizon_adaptive_2015}
    \item HIV \cite{fraser_virulence_2014}: \textbf{set-point viral load}
      correlated with transmission probability, rate of progression to AIDS
    \item data from Rakai, Uganda
    \item eco-evolutionary virulence dynamics: \cite{shirreff_transmission_2011}
  \end{itemize}

@@

\vspace{-1cm}

@@

gg_virtraj <- ggplot(ss,
                     aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model,linetype=model),lwd=1)+
    geom_ribbon(aes(fill = model), alpha=0.3)+
    labs(x="time (years)",y=Lspvl) +
    ## theme(axis.title.y = element_text(margin = margin(0,10,0,0)),
    ##  axis.title.x = element_text(margin = margin(10,0,0,0))) +
    facet_wrap(~model, nrow = 3) +
    zero_margin+
    theme(legend.text=element_text(size=16))
## scale_x_continuous(limits=c(0,2500))
print(gg_virtraj)
@

@@

@@

ggplot(mL3,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free",
               labeller = L)+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    ## leave a little extra room?
    scale_x_log10(expand=c(0,0.08),breaks=brkfun)+
    zero_margin
@               
              \end{block}
            \begin{block}{Conclusions}
\end{block}
\begin{block}{References}

>>>>

<<setup3,echo=FALSE,dev="tikz">>=
source("../R/hivFuns.R")
source("../R/Param.R")
ltab_mean <- apply(HIVpars_range, 1, geom_mean)
HIVpars.base <- as.HIVvirparlist(ltab_mean)
Lspvl <- "$\\log_{10}~\\textrm{SPVL}$"
pars <- expand(HIVpars.base)
dd <-with(pars,
     melt(data.frame(alpha=alpha,"transmission probability"=Beta2,
                     "duration (years)"=Duration2,check.names=FALSE),
          id.vars="alpha"))
gg_tradeoff <- ggplot(dd,aes(alpha,value))+facet_wrap(~variable,scale="free")+
    geom_line()+
    labs(x= Lspvl)
@

@@

\begin{block}{Summary}
\noindent
\hspace{-1.6cm}
Pathogens can evolve rapidly in response to changing
conditions (e.g., epidemic stage or public health interventions). 
Models of \textbf{eco-evolutionary dynamics} often
neglect important epidemiological processes, such as the 
dynamics of sexual partnerships. We
compared \textbf{models with a range of complexity} 
of partnership dynamics and extra-partnership contact.
\end{block}

@@

<<tradeoff_fig,echo=FALSE,fig.width=10>>=
print(gg_tradeoff)
@ 
  \begin{itemize}
    \item \textbf{tradeoff theory}: virulence evolution
      mediated by transmission-vs-clearance tradeoff
    \item still debated \cite{EbertBull2003,alizon_adaptive_2015}
    \item HIV \cite{fraser_virulence_2014}: \textbf{set-point viral load}
      correlated with transmission probability, rate of progression to AIDS (data from Rakai, Uganda)
    \item eco-evolutionary virulence dynamics: \cite{shirreff_transmission_2011}
  \end{itemize}

@@

@@

gg_virtraj <- ggplot(ss,
                     aes(tvec,mean,ymin=lwr,ymax=upr))+
    geom_line(aes(colour=model,linetype=model),lwd=1)+
    geom_ribbon(aes(fill = model), alpha=0.3)+
    labs(x="time (years)",y=Lspvl) +
    theme(axis.title.y = element_text(margin = margin(0,10,0,0)),
          axis.title.x = element_text(margin = margin(10,0,0,0))) +
    facet_wrap(~model, nrow = 3) +
    zero_margin+
    theme(legend.position="none")+
    scale_x_continuous(breaks=c(0,600))
print(gg_virtraj)
@

@@

\begin{itemize}

\end{itemize}

@@

ggplot(mL3,aes(LHSval,sumval,colour=model))+
    geom_point(pch=".",alpha=0.5)+
    facet_grid(sumvar~LHSvar,scales="free",
               labeller = L)+
    geom_smooth(se=FALSE)+labs(x="",y="")+
    ## leave a little extra room?
    scale_x_log10(expand=c(0,0.08),breaks=brkfun)+
    zero_margin
@               
              \end{block}

            \begin{block}{Conclusions}

\begin{itemize}
\item Random-mixing models gave the closest match to the most realistic models; extra-pair contact washed out the effects
of epidemiological structure
\item Large differences
in evolutionary dynamics among different epidemiological models
suggest caution in predicting evolutionary responses
\end{itemize}

\end{block}
\begin{block}{References}
